Estimating duration in the presence of misclassification

### 2024-02-08 biology-data.tex
Overleaf comments/edits
proof read ATACCC

SARS-CoV-2 is an RNA virus, but PCR amplifies DNA.
Therefore, before the PCR process can be used to detect SARS-CoV-2, the RNA must be converted to DNA.
This process is known as \emph{reverse transcription}.
In reverse transcription, a naturally-occurring enzyme known as reverse transcriptase generates cDNA (complementary DNA) from the RNA in the sample~\autocites{valasekPower}.
Any sequence of cDNA has a one-to-one correspondence with a sequence of RNA.
The cDNA can then be used in the standard PCR process.

@@

@@

The ATACCC study was a longitudinal study of contacts reported to the NHS Test and Trace contact tracing service~\autocite{hakkiOnset}.
Contacts were invited to participate if they were 5 years or older and  could be reached within 5 days of the index case's symptom onset date.
ATACCC was formed of two waves: ATACCC1 enrolled 393 contacts between 13th Sep 2020 and 31st Mar 2021 (when pre-Alpha and Alpha variants were circulating) and ATACCC2 enrolled 345 contacts between 24th May 2021 and 228th Oct 2021 (when the Delta variant was circulating).
Contacts were swabbed daily for up to 20 days, with a variety of tests performed on the swabs; relevant for this thesis is the RT-PCR test for SARS-CoV-2.
For all positive test results, the viral load was measured using the Ct value and then converted to a viral load using a standard curve.
Other data such as demography and symptoms was also collected.
For further details see \textcite{singanayagamDuration,hakkiOnset}.

@@

\subsubsection{Estimating positivity rate} \label{biology-data:sec:posivity}

>>>>

SARS-CoV-2 is an RNA virus, but PCR amplifies DNA.
Therefore, before the PCR process can be used to detect SARS-CoV-2, the RNA must be converted to DNA.
This process is known as \emph{reverse transcription}.
In reverse transcription, a naturally-occurring enzyme known as reverse transcriptase generates cDNA (complementary DNA) from the RNA in the sample~\autocite{valasekPower}.
Any sequence of cDNA has a one-to-one correspondence with a sequence of RNA.
The cDNA can then be used in the standard PCR process.

@@

\todo[inline]{Discuss noise in Ct values (batch effects, swabbing process, etc.) see commented-out text below}
% The noise arises from a variety of sources, including the sampling/swabbing process and batch effects in the RT-PCR process~\autocite{hanRTPCR}.
% A positive RT-PCR test has a Ct value.
% Ct values are linearly related to viral load (see \cref{E-biology-data:sec:PCR})
% A negative RT-PCR test signals either the viral load is too low to be detected, \ie below the limit of detection, or the test is a false negative.
% Positive RT-PCR tests, especially those with low viral loads, could be false positives, for example caused by contamination.

@@

The ATACCC study was a longitudinal study of contacts reported to the NHS Test and Trace contact tracing service~\autocite{hakkiOnset}.
Contacts were invited to participate if they were 5 years or older and  could be reached within 5 days of the index case's symptom onset date.
ATACCC was formed of two waves: ATACCC1 enrolled 393 contacts between 13th Sep 2020 and 31st Mar 2021 (when pre-Alpha and Alpha variants were circulating) and ATACCC2 enrolled 345 contacts between 24th May 2021 and 228th Oct 2021 (when the Delta variant was circulating).
Contacts were swabbed daily for up to 20 days, with a variety of tests performed on the swabs; relevant for this thesis is the RT-PCR test for SARS-CoV-2.
For all positive test results, the viral load was measured using the Ct value and then converted to a viral load using a standard curve.
Other data such as demography and symptoms was also collected, although I do not use this information.
For further details see \textcite{singanayagamDuration,hakkiOnset}.

@@

\subsubsection{Estimating positivity rate} \label{biology-data:sec:positivity}

### 2024-02-08 incidence-prevalence.tex
proof read ATACCC

A point process is a random function of continuous time $K(t)$.
$K(t)$ must take non-negative integer values and be monotonically increasing in $t$.
The interpretation of $K(t)$ is the number of events, here infections, that have occurred in the interval $(-\infty, t]$~\autocites[244]{yanDistribution}.
In the context of a disease, the point process can be discretised into the incidence on day $t$ as $Z_t = K(t) - K(t-1)$, where $t$ is an integer.
I will explain two approaches to modelling the infection process.

@@

A commonly-used point process for infectious disease epidemiology is the time-inhomogenous Poisson process~\autocites[e.g., in the context of HIV,][]{brookmeyerMethod}{paganoHIV}{rosenbergBackcalculation}{brookmeyerBackcalculation}.
% A time-homogenous Poisson process is the special case where $\nu(r)$ is constant.
Extensions to this model, which I do not consider here, have incorporated covariates, such as geography or age~\autocite[e.g.][]{diggleModeling}.

The first defining property of a Poisson process is that the counts are Poisson distributed.
Formally, $K(s) - K(t) \dist \Poi\left(\int^s_t \nu(r) dr \right)$ where $\nu(r) \geq 0$ is the \emph{intensity} of the infection process at time $r$~\autocites[244]{yanDistribution}.

>>>>

A point process is a random function of continuous time $K(t)$.
$K(t)$ must take non-negative integer values and be monotonically increasing in $t$.
The interpretation of $K(t)$ is the number of events, here infections, that have occurred in the interval $(-\infty, t]$~\autocite[244]{yanDistribution}.
In the context of a disease, the point process can be discretised into the incidence on day $t$ as $Z_t = K(t) - K(t-1)$, where $t$ is an integer.
I will explain two approaches to modelling the infection process.

@@

A commonly-used point process for infectious disease epidemiology is the time-inhomogenous Poisson process~\autocite[e.g., in the context of HIV,][]{brookmeyerMethod}{paganoHIV}{rosenbergBackcalculation}{brookmeyerBackcalculation}.
% A time-homogenous Poisson process is the special case where $\nu(r)$ is constant.
Extensions to this model, which I do not consider here, have incorporated covariates, such as geography or age~\autocite[e.g.][]{diggleModeling}.

The first defining property of a Poisson process is that the counts are Poisson distributed.
Formally, $K(s) - K(t) \dist \Poi\left(\int^s_t \nu(r) dr \right)$ where $\nu(r) \geq 0$ is the \emph{intensity} of the infection process at time $r$~\autocite[244]{yanDistribution}.
